### 2012-04-18 2.2.2.tex
chapters 2 and 4 have been swapped

% 2.2.2: Computational models

Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}. There is now a growing literature on computational cognitive models of phonotactic learning. By applying these models to the syllable contact data, it is possible to exhaustively search for statistically motivated constraints, elimianting the possibility of human error.

To evaluate these statistical models, they are first trained on the set of attested clusters, then put to the task of classifying all possible clusters into those which are and are not attested. Four metrics are used to evaluate these computational models. Classification accuracy corresponds to the probability of a classification being either a true positive ($t.p.$), correctly predicting a cluster to occur, or a true negative ($t.n.$), correctly predicting a cluster to be absent. Incorrect outcomes are either false positives ($f.p.$), incorrectly predicting a cluster to occur, and false negatives ($f.n.$), incorrectly predicting a cluster to be absent.

%\ex Binary classification contingency table: \\
%\begin{tabular}{c | l l} \toprule
%                             & \multicolumn{2}{c}{prediction} \\
%\midrule
%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
%                             & false negative ($fn$) & true negative ($tn$) \\
%\end{tabular}
%\begin{tabular}{|l l|}
%\toprule
%true positive ($tp$)  & false positive ($fp$) \\
%false negative ($fn$) & true negative ($tn$) \\
%\bottomrule
%\end{tabular}
%\xe

\ex $\displaystyle \textrm{Accuracy} = \frac{t.p. + t.n.}{t.p. + t.n. + f.p. + f.n.}$ \xe

\noindent
A few other metrics can be used to assess classification performance. Precision corresponds to the probability that a cluster predicted to occur is observed in the data.

\ex $\displaystyle \textrm{Precision} = \frac{t.p.}{t.p. + f.p.}$ \xe

\noindent
Recall (also known as sensitivity) is the probability a observed cluster is predicted to occur.

\ex $\displaystyle \textrm{Recall} = \frac{t.p.}{t.p. + f.n.}$ \xe

It is possible to maximize recall at the expense of precision, by predicting more clusters to be attested, or to increase precision at the expense of recall by predicting fewer clusters to be attested. A common way to balance these concerns is the $F_1$ score, which is the harmonic mean of precision and recall.

\ex $\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$ \xe

\subsubsection{Null baseline}

The 840 possible clusters have an 18.7\% of saturdaion rate. As a consequence of this sparsity, a null model which classifies all clusters as unattested achieves 81.3\% accuracy without the need to posit any constraints either derived or static. Since this model does not have any positive classifications, either true or false, precision, recall, and $F_1$ are undefined.

\subsubsection{Alternations}

As was shown above, \textsc{Obstruent Voice Assimiation}, \textsc{Coda Nasal Place Assimilation}, and \textsc{Degemination} are reliable predictors of what clusters are and are not attested and have few lexical exceptions. I generate another baseline by predicting only those clusters which do not violate any of these generalizations to be attested. This provides a significant improvement to the null baseline (accuracy = 0.857). As precision (0.401) is smaller than recall (0.708), there are more false positives (i.e., unattested clusters that do not violate any of these three alternation-derived generalizations) than there are false negatives (i.e., attested exceptions to these derived generalizations).

\subsubsection{\citet{Pierrehumbert1994}}

\citet{Pierrehumbert1994} first proposes to measure the well-formedness of syllable contact clusters by the independent probabilities of codas and onsets, and \citet{Coleman1997} extend this model to whole words. \citet{Pierrehumbert1994} uses both medial and initial/final codas and onsets to compute these probabilities, but I obtain a better fit to the data by only using medial coda and onset frequencies. Medial coda and onset probabilities are multipled to obtain a well-formedness score. This produces a continuous value between 0 and 1. To derive a binary classification from these values, I use a soft-margin support vector machine \citep{Cortes1995} with a linear kernel. This technique derives a ``decision stump'' providing the best single split into attested and an unattested clusters. This is not put forth as a cognitively plausible model for relating this expected probability to attestation: it is simply the upper bound that such a model could reach.

Unfortunately, this score only provides a small improvement to the null baseline
 (accuracy = 0.839), and has a lower accuracy and $F_1$ (0.328) than the alternation baseline.

\subsubsection{\citet{Hayes2008a}}

\citeauthor{Hayes2008a} develop a sophisticated method of learning phonotactic distributions from positive data. They develop software for estimating probability distributions over sequences of (sets of) phonological features using the principle of maximum entropy and making few assumptions about the independence of features. As above, a support vector machine with a linear kernel is used to turn these probability values into binary predictions. The software provided by these authors exposes many ``metaparameters'', operational decisions to be made in conducting an experiment with this system. To avoid any potential bias induced by the choice of these metaparameters, I use the same settings as used by \citet{Hayes2008a} in their similar study of English onset consonants. This includes the ``accuracy schedule'' [0.001, 0.01, 0.1, 0.2, 0.3] and no constraints on the number of constraints learned, as well English consonantal features used by \citeauthor{Hayes2008a}. Since this model is inherently stochastic, producing slightly different outcomes on each run, I follow the practice of \citeauthor{Hayes2008a} (p.~396) and report the lowest-accuracy of ten runs, though in general there is not a great deal of variation between individual runs.

This model provides a small improvement in accuracy (0.877) and $F_1$ (0.703) compared to the alternation baseline, but it has lower recall (0.642) than this baseline. This indicates that the model produces many false negatives.
%; it assigns the lowest probability value to 122 unattested clusters.
This seems to further indicate that this model internalizes narrower phonotactic generalizations than those derived from the three alternation-derived constraints; for instance, it assigns the highest probability to unatteested *[m.kl], which is ruled out by \textsc{Coda Nasal Place Assimilation}. On the other hand, it also rules out clusters like those in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are consistent with all three alternation-derived constraints.

\citet{Berent2012} identify an important defect with the publically-available version of the \citeauthor{Hayes2008a} model used here: it cannot directly incorporate constraints like the alteration-derived constraint disfavoring geminate consonants, because it does not include variables over phonological features. However, the model appears to have experienced no problems internalizing the generalization that geminate clusters are unattested.

The results for the four models are summarized below.

\ex Summary of computational model accuracy: \vspace{6pt} \\
\begin{tabular}{l | r r r r}
\toprule
                          & accuracy & precision & recall & $F_1$ \\ 
\midrule
Null baseline             & 0.813    & n.a.      & n.a.   & n.a.  \\
Alternations              & 0.857    & 0.401     & 0.708  & 0.512 \\
\citet{Pierrehumbert1994} & 0.839    & 0.210     & 0.750  & 0.328 \\
\citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
\bottomrule
\end{tabular}
\xe

\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{I am grateful to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are more frequent than those like [t.j], with rising sonority. \citeauthor{McGowan2011} reports, however, that this accounts for only a tiny amount of the variance in frequency ($R^2 = 0.077$). Change in sonority provides no improvement to the null baseline, and so this model is not reported here.

>>>>

\subsection{Assessing models}

\subsubsection{Basic correlations}
\subsubsection{Intramodal correlations}
\subsubsection{Residual correlations}

### 2012-04-18 2.3.tex
Rearranged and some writing on 2-5.
Drastic restructuring and some writing on chapters 2-3.
More arranging.
chapters 2 and 4 have been swapped

% 2.3: Conclusion

The above is a stark reminder.

A reasonable objection to the arguments presented in this chapter is to view the results as little more than an indictment of the results of \citealt{Pierrehumbert1994}.

However, the fact that state-of-the-art models are not capable of providing large improvements to the predictive accuracy indicates

 do not have a statistically significant effect on the shape of the English lexicon, but that experiments might turn up evidence that speakers have internalized

shown to be aware of static constraints if they reached statistical significance.

 statistically reliable static constraints could be identified

The of

Regarding the historical developments,

\citet{Martin2007}

On the other hand, patterns created by sound change are not guaranteed to persist over time. 
One example of non-persistence is discussed by \citet{Iverson2005}.  
Around 1100 CE, Old English \emph{sk} became [ʃ]. 
This sound change introduced no alternations.
Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{V\lm sk\#} words when the
 change was actuated, and \emph{V\lm sh\#} continues to be rare in Modern English. 
What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.

A third pattern is that a historically inherited pattern

\citet[][140]{Frisch2004} suggest that the strong tendency for the first and second consonants of the Arabic root to be non-identic
al is the ``a diachronic result of a processing constraint that disfavors repetition.'' 
Unfortunately, there is no evidence that this pattern is diachronic other than in the sense that it appears to be inherited from the proto-language: there is simply no Proto-Semitic verb roots with identical first and second consonants \citep[][178]{Greenberg1950}. 
In other Semitic languages, the inherited patern has experienced considerable erosion.

\ex Tigrinya roots with identical first and second consonants \citep{Buckley1990a}: \\
\begin{tabular}{l l l l}
a. & lʌlʌw     & `scorch'                   & (< Ge'ez \emph{lʌwlʌw} `inflame')     \\
   & mʌmʌy     & `winnow'                   & (< Ge'ez \emph{mʌymʌy} `distinguish') \\
%   & mʌmʌt & `pick out loot' & (< 
b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
\end{tabular}
\xe

Similar exceptions are found in 
%Amharic (\citealp[][?]{Broselow1984}, \citealp[][?]{McCarthy1985}) and 
Hebrew \citep[][29]{Bat-El2005}.

The next two chapters return to the question of synchrony, addressing the relationship between statistical patterns in the lexicon and speakers' behaviors when presented with underrepresented sequences.

>>>>

\section{Conclusions}

It is not the case that this evidence is inconsistent with the hypothesis that there are additional degrees of wordlikeness: it is just a demonstration that these data to not require it.

### 2012-04-18 2.tex
Drastic restructuring and some writing on chapters 2-3.
More arranging.
chapters 2 and 4 have been swapped

% 2: Accidental and structural gaps in English syllable contact

In the preceding chapter, it was proposed that

Reconsider the null hypothesis presented in the preceding chapter:

\ex The null hypothesis: \\
Universal grammar does not countenance constraints on the contents of URs
\xe

The idea

relationship to proposal

One primary way which this null hypothesis has been critiqued is by demonstrating that the lexicon of some language 
exhibits dispreferences or gaps corresponding to.

argue by trends in the lexicon

possibility of sparsity (C\&H, Fischer/Vogt and recent stats)

Of course, to obtain the desired results, we must guarantee that each sequence structure rule reflects a general systematic fact about the language, and not a fact which is due merely to the existence of accidental gaps in the lexicon. \citep[][401, fn.~8]{Stanley1967}

In the preceding chapter, it was proposed that

Lexicon

This appears to be exceptionlessly true of Arabic

That even gradient patterns

One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. 
Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a} but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Dmitrieva2008a,Dmitrieva2008b,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.

The absence of 
The claim here is that however stark the absence or underrepresentation of some form is, it

This hypothesis admits the possibility that there may be accidental gaps in the lexicon, a possibility which predates generative thinking:

\begin{quotation}
\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language. \citep[][16]{Fischer-Jorgensen1952}
\end{quotation}

\noindent
Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters

\begin{quotation}
Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters. The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
\end{quotation}

\citet{Chomsky1965}

This chaper demonstrates that apparently systematic gaps may arise even in the absence of phonotactic preferences, and that this possibility compromises attempts to show that such preferences shape the lexicon.

>>>>

\chapter{Gradient and categorical aspects of wordlikeness judgements}
\label{gradience}

(though few if any phonologists regard $n$-gram segment probabilities as a core component of phonology)

%\begin{quotation}
%``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
%\end{quotation}

\input{2.1}
\input{2.2}
\input{2.3}

### 2012-04-18 3.1.1.tex
chapters 2 and 4 have been swapped

% 3.1.1: Alternation vs. phonotactic

\subsubsection{Roundness harmony}

\subsubsection{Labial attraction}

\citet{Zuraw2000}
\citet{Clements1982}
\citet{Inkelas1997}

>>>>

\subsection{Backness harmony}

One of the most salient properties of Turkish is its harmony system, which \citeauthor{Lees1966b} (\citeyear[][35]{Lees1966b}, \citeyear[][284]{Lees1966a}) derives by means of feature spreading rules. The least constrained of these processes is the rule of backness harmony described in this section.

\subsubsection{Phonological description}

Backness harmony spreads the value of \textsc{Back} rightward over any intervening consonants onto the next vowel.

%\citep[after][229]{Clements1982}: \\
\begin{example}
\textsc{Backness Harmony}:

\xymatrix@R=24pt@C=24pt{
\txt{V}                                        & \txt{C$_0$} & \txt{V} \\
\txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{..}[urr] \\
}
\end{example}

Along with the rule of \textsc{Roundness Harmony} discussed in the next section, this rule accounts for a great deal of the suffix allomorphy found in Turkish. The nominative plural (nom.pl.) is formed by adding the suffixes \emph{-ler} or \emph{-lar} to the nominative singular (nom.sg.).

\begin{example}
The Turkish nominative:

\vspace{0.5\baselineskip}
\begin{tabular}{l l l l l}
   & \emph{nom.sg.} & \emph{nom.pl.} \\ 
a. & ip             & ipler          & `rope' & \citep[][216]{Clements1982} \\
   & el             & eller          & `hand'    \\
   & köy            & köyler         & `village' \\
   & yüz            & yüzler         & `face'    \\
   & kız            & kizlar         & `girl'    \\
   & sap            & saplar         & `stalk'   \\
   & son            & sonlar         & `end'     \\
   & pul            & pullar         & `stamp'   \\
b. & neden          & nedenler       & `reason'  & (TELL) \\
   & boğaz          & boğazlar       & `throat'  \\
   & kiler          & kilerler       & `pantry'  \\
   & pelür          & pelürler       & `tissue paper' \\
   & sapık          & sapıklar       & `pervert' \\
\end{tabular}
\end{example}

\noindent
\textsc{Backness Harmony} correctly predicts the choice of suffix. There are a few complications, however. First, while most polysyllabic roots conform to \textsc{Backness Harmony} (\lastx b), not all do. In this case, suffix vowels generally harmonize with the final root vowel (\nextx a). Yet there is also a small class of nouns which exhibit [$-$\textsc{Back}] suffix vowels despite the fact that their final root vowel is [$+$\textsc{Back}] (\nextx b).

\begin{example}
Exceptional Turkish nominatives:

\vspace{0.5\baselineskip}
\begin{tabular}{l l l l l}
   & \emph{nom.sg.} & \emph{nom.pl.} \\
a. & mezar          & mezarlar       & `grave' & \citep{TELL} \\
   & model          & modeller       & `model' \\
   & silah          & silahlar       & `weapon'     \\
   & memur          & memurlar       & `bureaucrat' \\
   & sabun          & sabunlar       & `soap'       \\
b. & saat           & saatler        & `hour, clock' \\
   & harf           & harfler        & `(alphabetic) letter' \\ %& \citep{Goksel2005}
   & etol           & etoller        & `fur stole' \\
\end{tabular}
\end{example}

While it is uncontroversial that disharmonic suffixes (\lastx b) are no more than sporadic exceptions to \textsc{Backness Harmony}, the status of root disharmony (\lastx a) has been the subject of much debate. \citet[][212, 289]{Anderson1974} views the fact that disharmonic roots still trigger suffix harmony as evidence that suffix harmony is distinct from root harmony, the latter being a sequence structure constraint.

%Once additional source of evidence on root (dis)harmony is inconclusive. There is a small class of bisyllabic words in which the second vowel, always [$+$\textsc{High}, $-$\textsc{Back}], alternates with zero.

%\ex High-vowel/zero alternations \citep[][243]{Clements1982}: \\
%\begin{tabular}{l l l l}
%   & nom.sg. & gen.sg. \\
%a. & fikir   & fikri  & `idea' \\
%   & hüküm   & hükmün & `judgement' \\
%%  & filim   & filmi & `film' & \citep[][178]{Inkelas2001} \\
%b. & vakit   & vaktin & `time' \\
%   & rahim   & rahmin & `womb' \\
%\end{tabular} \xe
%
%\noindent
%It is possible that \textsc{Backness Harmony} might produce a fluctuating \emph{ı} after root \emph{a}, but this does not obtain (\lastx b). However, this might simply indicate that the fluctuating vowel is epenthetic and that harmony applies before epenthesis (see \citealt{Clements1982} for both sides of this argument), making it less than a counterexample.

Underspecification provides an alternative under which the exceptionality of disharmonic roots is only apparent. While it is clear that the first vowel of Turkish roots are contrastively specified for backness (e.g., \emph{kül} `ash' vs.  \emph{kul} `servant', \emph{kepek} `bran' vs. \emph{kapak} `lid'), there are also some minimal pairs with regards to backness (dis)harmony (e.g., \emph{deve} `camel' vs. \emph{deva} `medicine', \emph{sene} `year' vs. \emph{sena} `praise'). This supports the possibility, first suggested by \citet{Clements1982}, that harmonic roots may be underspecified and disharmonic roots fully specified, exemplified below.

%There is some further evidence that individual vowels may differ in specification for this feature even within individual roots or affixes. For instance, the present continuous suffix has harmony-determined allomorphs \emph{-iyor}, \emph{-üyor}, \emph{-ıyor}, \emph{-uyor}, but the \emph{o} of the suffix is invariant. A similar situation might obtain in Turkish roots.

\begin{example}
Underlying specification of \textsc{Back} for (dis)harmonic roots:

\xymatrix@R=24pt@C=24pt{
\txt{a.} & \txt{s} & \txt{V} & \txt{n} & \txt{V}  & \txt{\emph{sene} `year'} \\
&        & \txt{[$-$\textsc{Back}]}\ar@{-}[u]     \\
\txt{b.} & \txt{s} & \txt{V} & \txt{n} & \txt{V} & \txt{\emph{sena} `praise'} \\
         &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u] & & \txt{[$+$\textsc{Back}]}\ar@{-}[u]
}
\end{example}

\noindent
Harmonizing suffix vowels will also be underspecified for \textsc{Back}. Of course, \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being the use of a a \textsc{Structure Preservation} condition \citep{Kiparsky1985}. Unfortunately, this detail reintroduces the duplication alluded to above. Any condition on the rule of \textsc{Backness Harmony} which prevents overwriting entails that it has no control over the distribution of harmonic and disharmonic roots, and thus fails to account for predominance of harmonic roots.

However, the theory of exceptionality presented in \emph{SPE} (p.~374f.) provides a direct account of suffix harmony in disharmonic roots without duplication. In \emph{SPE}, the specification of the target (i.e., the segment to be changed) of every rule $R$ must be [$+$rule $R$] by convention.  A root or affix which fails to undergo $R$ despite otherwise matching the structural description is simply said to be marked [$-$rule $R$], and thus failing to meet the full structural description. If disharmonic roots are [$-$\textsc{Backness Harmony}], then suffix vowels are correctly predicted to undergo \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the environment, which is not required to be [$+$rule $R$].

\subsubsection{Psycholinguistic evidence}

The evidence from alternations leaves open the question of whether speakers internalize a generalization regarding the tendency of roots to conform to \textsc{Backness Harmony}. One piece of evidence that speaks in favor of root-internal harmony comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but it corresponds to a morphological process native to the related language Tuvan, in which it conveys a sense of ``vagueness or jocularity'', and \citeauthor{Harrison2001} report that it can be quickly taught to even young Turkish speakers. The game consists of reduplication of the base and replacement of the first vowel in the reduplicant with \emph{a} or \emph{u}.

In both Tuvan and Turkish, the unchanged reduplicant vowel is also affected. Reduplication interacts with root harmony in both Tuvan and Turkish. When the base is harmonically [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] results in what \citeauthor{Harrison2001} call ``reharmonization'' (\nextx a).

\begin{example}
\label{redupgame}
Turkish reduplication game \citep[][231]{Harrison2001}:

\vspace{0.5\baselineskip}
\begin{tabular}{l l l l}
a. & kibrit & kibrit-kabrıt & `match'    \\
   & bütün  & bütün-batın   & `whole'    \\
b. & mali   & mali-muli     & `Mali'     \\
   & butik  & butik-batik   & `boutique' \\
\end{tabular}
\end{example}

\noindent
Under the underspecification hypothesis, non-initial harmonic vowels lack an underlying \textsc{Back} feature, so it comes as no surprise that changing the \textsc{Back} specification of the root-initial vowel results in reharmonization. And the full specification of disharmonic roots correctly predicts that they will be exempt from reharmonization (\lastx b), which is also borne out in Tuvan and in an unrelated Finnish language game \citep{Campbell1986}.

A number of studies have found that speakers of Finnish use disharmony in speech processing experiments.\footnote{Thanks to Charles Yang for pointing out the relevance of these studies to me.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which are harmonic for the feature \textsc{Back}. These stimuli are auditorily presented to subjects who are asked to press a button when the nonce trisyllable ends with a target nonce disyllable, or a real disyllabic word. Speakers are quicker to press the button when the prefix and disyllabic word disagree for \textsc{Back}. These results suggest that speakers are attuned to the fact that disharmonic transisitions are good predictors of word boundaries. If speakers have also internalized the converse generalization, that harmonic transistions are more likely to be root-internal, then there is additional evidence that harmony is active not just in Finnish affix alternations but also in roots.

\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect on word-spotting as it does in Finnish: speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by the pseudo-prefix \emph{gölü-}, a disharmonic transistion, than when it is preceded by the pseudo-prefix \emph{golu-}, a harmonic transition. \citeauthor{Kabak2010} find that effect of harmony does not obtain for speakers of French, a language which lacks vowel harmony. As in Finnish, the results imply speakers have internalized the predominance of root-internal harmony.

It seems that the root-internal harmony bias is in fact learned by Turkish speakers very early. The pseudoword spotting experiment has been adapted for 9-month-old Turkish infants by \citet{Kampen2008}. Infants are familiarized with harmonic disyllabic pseudowords bearing a pseudo-prefix, which may be harmonic or disharmonic. At test time, the infants are played the disyllabic pseudowords in isolation using the head turn preference paradigm. Infants show a preference to listen to those pseudowords which were familiarized with a disharmonic pseudo-prefix over those which were familiarized with a harmonic pseudo-prefix. This preference is not observed in 9-month-old infants learning German, which also lacks vowel harmony. Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants show no such preference.

### 2012-04-18 3.1.2.tex
chapters 2 and 4 have been swapped

% 3.1.2: Lexical statistics

This is essentially the hypothesis pursued in the previous chapter.

\citet{TELL}
\citet{Inkelas2001}

%We might, for instance, consider methods based not on statistical \citep{Yang2005,Gorman2011b}.

>>>>

\subsection{Roundness Harmony}

The rule of roundness harmony and the data that motivates it overlaps with the preceding evidence for \textsc{Backness Harmony}.

\subsubsection{Phonological description}

The environment for roundness harmony differs from \textsc{Backness Harmony} in that it requiers the target to be [$+$\textsc{High}].

\begin{example}
\textsc{Roundness Harmony}:

\xymatrix@R=24pt@C=24pt{
\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{..}[drr] &             & \txt{[$+$\textsc{High}]} \\
\txt{V}                                         & \txt{C$_0$} & \txt{V}\ar@{-}[u] \\
}
\end{example}

This rule, in concert with \textsc{Backness Harmony}, accounts for the forms of the dative singular (dat.sg.) and genitive singular (gen.sg.), among other suffixes.

\begin{example}
Turkish nominal suffix allomorphy:

\vspace{0.5\baselineskip}
\begin{tabular}{l l l l l l l}
   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
a. & ip             & ipi            & ipin           & `rope' & (CS:216) \\
   & el             & eli            & elin           & `hand'    \\
   & kız            & kızı           & kızın          & `girl'    \\
   & sap            & sapı           & sapın          & `stalk'   \\
   & yüz            & yüzü           & yüzün          & `face'    \\
   & köy            & köyü           & köyün          & `village' \\
   & pul            & pulu           & pulun          & `stamp'   \\
   & son            & sonu           & sonun          & `end'     \\
b. & boğaz          & boğazı         & boğazın        & `throat'  & (TELL) \\
   & pelür          & pelürü         & pelürün        & `tissue paper' \\
   & döviz          & dövizi         & dövizin        & `currency' \\
   & yamuk          & yamuğu         & yamuğun        & `trapezoid' \\
   & ümit           & ümiti          & ümitin         & `hope'     \\
\end{tabular}
\end{example}

\noindent
Much as was the case for \textsc{Backness Harmony}, there are disharmonic roots which participate in suffix harmony. Once again, underspecification of harmonic roots and full specification of disharmonic roots allows for the use of the \emph{SPE} exception convention.

\subsubsection{Psycholinguistic evidence}

The only psycholinguistic evidence for \textsc{Roundness Harmony} comes from \citeauthor{Harrison2001}'s language game (\ref{redupgame}). As shown above, the second vowel in the reduplicated form of \emph{bütün} `whole' is \emph{bütün-batın}. The second vowel in the base, \emph{ü}, is reharmonized as [$+$\textsc{Back}, $-$\textsc{Round}], indicating that \textsc{Roundness Harmony} also participates in reharmonization.

### 2012-04-18 3.1.tex
chapters 2 and 4 have been swapped

% 3.1: Wordlikeness predictions

\citet{Zimmer1969}

>>>>

\section{Turkish vowel sequence structure constraints}

\citet{Lees1966a,Lees1966b} proposes three sequence structure constraints on Turkish vowels, and these constraints have been discussed extensively in the subsequent literature. Below, these constraints are formalized and related to Turkish alternation phonology and various sources of external evidence. A traditional binary feature specification for the Turkish vowel system, given below, is assumed without discussion.

\begin{example}
Turkish vowel features:

\vspace{0.5\baselineskip}
\begin{tabular}{c | c c c c}
                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
\buf[$+$\textsc{High}] & i & ü [y] & ı [ɯ] & u \\
\buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
\end{tabular}
\end{example}

\input{3.1.1}
\input{3.1.2}
\input{3.1.3}

### 2012-04-18 3.2.2.tex
chapters 2 and 4 have been swapped

\begin{quotation}
Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research. \citep[][196, emphasis in original]{Inkelas2001}
\end{quotation}

@@

\ex $\displaystyle \gamma = \frac{C - D}{C + D}$ \xe

@@

Backness harmony wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
\midrule
pemez & 30                   & pemaz & 2  \\
teriz & 28                   & terız & 3  \\
tokaz & 26                   & tokez & 6  \\
tipez & 24                   & tipaz & 8  \\ 
terüz & 19                   & teruz & 13 \\ % roundness violator
\bottomrule
\end{tabular}

@@

Labial attraction wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{aPu} & \multicolumn{2}{l}{aPı} \\
\midrule
tafuz & 21              & tafız & 11 \\
pamuz & 17              & pamız & 15 \\
tapuz & 17              & tapız & 15 \\
mavuz & 16              & mavız & 16 \\
tabuz & 16              & tabız & 16 \\
\bottomrule
\end{tabular}

>>>>

\begin{quote}
Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research. \citep[][196, emphasis in original]{Inkelas2001}
\end{quote}

@@

\begin{example}
$\displaystyle \gamma = \frac{C - D}{C + D}$
\end{example}

@@

\begin{example}
Backness harmony wordlikeness forced choices \citep[314]{Zimmer1969}:

\vspace{0.5\baselineskip}
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
\midrule
pemez & 30                   & pemaz & 2  \\
tipez & 24                   & tipaz & 8  \\ 
terüz & 19                   & teruz & 13 \\ % roundness violator
teriz & 28                   & terız & 3  \\
tokaz & 26                   & tokez & 6  \\ % roundness violator
\bottomrule
\end{tabular}
\end{example}

\subsubsection{Roundness harmony}

\begin{example}
Roundness harmony forced choice wordlikeness judgements:

\vspace{0.5\baselineskip}
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
\midrule
pörüz & 32 & pöriz & 0  \\
tatız & 20 & tatuz & 12 \\
tüpüz & 31 & tüpiz & 1  \\
takız & 22 & takuz & 10 \\
\bottomrule
\end{tabular}
\end{example}

%tamaz & 26 & tamoz & 6  \\
%putoz & 25 & putaz & 7  \\

@@

\begin{example}
Labial attraction wordlikeness forced choices \citep[314]{Zimmer1969}:

\vspace{0.5\baselineskip}
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{aPu} & \multicolumn{2}{l}{aPı} \\
\midrule
pamuz & 17              & pamız & 15 \\
tafuz & 21              & tafız & 11 \\
tapuz & 17              & tapız & 15 \\
mavuz & 16              & mavız & 16 \\
tabuz & 16              & tabız & 16 \\
\bottomrule
\end{tabular}
\end{example}

### 2012-04-18 3.tex
chapters 2 and 4 have been swapped

This is not a universal, but rather something English speakers must learn: some languages, such as Moroccan Arabic, do in fact permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet').

%bni 'building'
%bnin 'delicious'
%bniqa 'closet' 
%bnat 'girl' 
%bnita 'girl (dim.)'

a matter of universal

In their illustration of

One of the

One of the most compelling findings

@@

@@

\citet{Daland2011}

>>>>

glossematics

\citet{Chomsky1965} propose

In retrospect, it seems somewhat odd that \citeauthor{Chomsky1965} would coh
It might 
In defining the

@@

This is not a universal, but rather something English speakers must learn, as some languages, such as Moroccan Arabic, do in fact permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet');

\citet{Lees1966b,Lees1966a}

@@
